---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# GIFTr

<!-- badges: start -->
<!-- badges: end -->

GIFTr package in intended to help course creators to upload questions to MOODLE and other LMS for quizzes. GIFTr takes dataframe or tibble of questions of four types: mcq, numerical questions, true or false and short answer,and export it a text file formatted in MOODLe GIFT format. You can prepare a spreadsheet in any software and import it in R and generate any number of questions with HTML, markdown and LATEX support.

## Installation


from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("omarelashkar/GIFTr")
```
## Example


```{r example}
library(GIFTr)
# load GIFTrData
data("GIFTrData")

# See the structure of the data
head(GIFTrData , 2)
str(GIFTrData)

#Create quiz.txt file with GIFT format 
GIFTr::GIFTr(data = GIFTrData, questions = 3, answers = c(4:8), 
             categories = 1, question_type = 9, output = 'quiz.txt')
```

The generated file can be imported easily by MOODLE or any other LMS that supports GIFT.

For further documenation instructions, check the package documentation.
