---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# GIFTr

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/omarelashkar/GIFTr.svg?branch=master)](https://travis-ci.org/omarelashkar/GIFTr)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/omarelashkar/GIFTr?branch=master&svg=true)](https://ci.appveyor.com/project/omarelashkar/GIFTr)
[![DOI](https://zenodo.org/badge/208570969.svg)](https://zenodo.org/badge/latestdoi/208570969)
<!-- badges: end -->

'GIFTr' package in intended to help course creators to upload questions to MOODLE and other LMS for quizzes. 'GIFTr' takes dataframe or tibble of questions of four types: mcq, numerical questions, true or false and short answer,and export it a text file formatted in MOODLE GIFT format. You can prepare a spreadsheet in any software and import it in R and generate any number of questions with HTML, markdown and LATEX support.

## Installation


from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("omarelashkar/GIFTr" , build_vignettes = TRUE)
```
## Example


```{r example}
library(GIFTr)
# load GIFTrData
data("GIFTrData")

# See the structure of the data
head(GIFTrData , 2)
str(GIFTrData)

#Create quiz.txt file with GIFT format 
GIFTr::GIFTr(data = GIFTrData, questions = 3, answers = c(4:8), 
             categories = 1, question_type = 9, output = 'quiz.txt')
```

```{r message=FALSE, warning=FALSE, include=FALSE}
file.remove("quiz.txt")
```

The generated file can be imported easily by MOODLE or any other LMS that supports GIFT.

For further documentation instructions, check the package vignette and documentation.

```r
#open vingette
vignette("GIFTr_tutorial")
```
